\documentclass[lang=en, hanging-titles=true]{skrapport}

\usepackage[backend=biber]{biblatex}
\addbibresource{References.bib}

\usepackage[hidelinks]{hyperref}
\usepackage{graphicx} % allow embedded images
  \setkeys{Gin}{width=\linewidth,totalheight=\textheight,keepaspectratio}
  \graphicspath{{../figs/}} % set of paths to search for images
\usepackage{listings}
\usepackage{listings-rust}
\usepackage{ragged2e}
\usepackage{ctable}
\usepackage{cleveref}
\usepackage{multicol} % multiple column layout facilities

\raggedright
\colortheme{skdoc}
\title{An arm build on Promises}
\author[dskleingeld@gmail.com]{David Kleingeld}

% notes: 
% https://www.rapitasystems.com/blog/what-really-happened-software-mars-pathfinder-spacecraft

\begin{document}

\maketitle
\tableofcontents

\section{Introduction}
Recent years have seen great strides in the field of humanoid robotics. As we get closer to interactions between robots and machines we will need to guarentee safe operations. For this the movement of the robot needs to be strictly controled. Classicly robots are controlled using software build on \textit{real time operating systems} that guarentee strict timing between operations. This at the cost of limitations and complexity. This complexity offers many opportunities for unsafe mistakes. Here I try an alternative, controlling a robot arm without threads, OS, heap and all but excluding the chance for undefined behaviour. This while using easy to reason about software primitives. The arm itself is very simple, constructed out of lego and limited to planar movement. Control is done by a 64kB microcontroller.

In the next section I will introduce the \textsc{Async} software primitive I use instead of threads to do 'concurrent' work. Then in \cref{sec:impl} I will describe my implementation before discussing how this new approach performs in \cref{sec:results}. Finally I will conclude if I succeeded building a simple robotic arm this way and wheather this approach merits further study.

\section{Async}
Before I can describe my implementation I need to introduce \textsc{Async} a syntactic language feature that allows for easy constuction of asynchronous non blocking functions. \textit{Asynchronous} programming lets us write concurrent, not parallel, tasks while looking awfully similar to normal blocking pragramming. It is a good alternative to \textit{event-driven} programming which tends to be verbose and hard to follow. All \textsc{Async} systems are build around special function that do not return a value but rather a \textit{promise} of a \textit{future} value. When we need the value we tell the program not to continue untill the promise is fulfilled. Lets lok at the example of downloading 2 files:

\begin{lstlisting}[language=rust, style=boxed, tabsize=2]
async fn get_two_sites_async() {
	// Create two different "futures" which, when run to 
	// completion, will asynchronously download the webpages.
	let future_one = download_async("https://www.foo.com");
	let future_two = download_async("https://www.bar.com");

	// Run both futures to completion at the same time.
	let futures_joined = join!(future_one, future_two);
	// Run them to completion returning their return values
	let (foo, bar) = futures_joined.await;
	some_function_using(foo,bar);
}
\end{lstlisting}

Notice the \texttt{async} keyword in front of the function definition it means the function will return a promise to complete in the future. The \texttt{join!} statement on line 8 combines the two promises for a future awnser to a single promise for two awnsers. In line 10 we await or 'block' the program until \texttt{futures\_joined} turns into two value. Those can then be used in normal and async functions.

The caller of our \texttt{async} \textit{get\_two\_sites\_async} function will need to be an another async function that can await \textit{get\_two\_sites\_async}, or it can be an executor. An executor allows a normal function to await async functions.

Lets go through our example again explaining how this mechanism could work. The syntax and workings of async differ a lot here we will look at the language \textit{Rust}. In rust these promises for a future value are called futures. Until the program reaches line 10 no work on downloading the example sites is done. This is not a problem as the results, \textit{foo} and \textit{bar}, are not used before line 11. The runtime will start out working on downloading \texttt{www.foo.com}, probably by sending out a dns request. As soon as the dns request has been send we need to wait for the awnser, we need it to know to which ip to connect to download the site. At this point the runtime will instead of waiting start work on downloading bar where it will run into the same problem. If by now we have recieved an awnser on our dns request for \textit{www.foo.com} the runtime will continue its work on downloading foo. If not the runtime might continue on some other future availible to it that can do work at this point.

\clearpage
\section{Implementation}
\label{sec:impl}
Here I will discuss how I used async to get rid of threads and an OS, why and how my implementation was written heapless, how I all but excluding the chance of undefined behaviour and finally the hardware used.
% TODO mention double pid loop

\subsection{Threadless concurrency}
By combining the described async with an endless loop we can maintain multiple perhipherals. The key trick is the composability of futures using \textit{join} and a function called \textit{select}. My robot arm consists of three motor that need to be maintained, that is they must hold or move to a certain position. To achieve this each of the motors has a \texttt{async} function \textit{maintain} that controls the speed of the motor by regulating its power. It looks a little %
\footnote{I simplified the rust syntax, see the function \texttt{maintain} in \texttt{motor.rs} \href{https://github.com/dskleingeld/robotic-arm/blob/main/src/hinge/motor.rs}{online}} 
like this:

\begin{minipage}{\linewidth} % do not break 
\begin{lstlisting}[language=rust, style=boxed, tabsize=2]
async fn maintain(&mut self) {
	loop {
		select! {
			encoder => self.state.update(dist, spd),
			changed => {
				let speed = self.controls_get_speed()
				self.pid.set_target(speed)
			},
			timeout => self.state.update(0,0),
		}
	}
	let power = self.pid.update(self.state.speed)
	self.driver.set(power)
}
\end{lstlisting}
\end{minipage}

The \texttt{loop} on line 2 makes the function run forever. On line 3 \texttt{select!} is a special async function with similar syntax to match statement\footnote{switch/case statement in C}. The \textit{arms} of the select, here \texttt{encoder}, \texttt{changed} and \texttt{timeout}, are futures. The select function concurrently awaits its futures. As soon as one of the futures resolves to a value the code after the arrow (=>) of that future will run. No more work will happen on the other futures. That means the code above either:
\begin{enumerate}
	\item Updates the state of the motor
	\item Sets the motors target (pid) target position
	\item Update the state of the motor to reflect standing still
\end{enumerate}
Then the new power for the motor is calculated and set in lines 12 and 13.

To control multiple motors simply join on their maintain futures: 

\begin{minipage}{\linewidth} % do not break 
\begin{lstlisting}[language=rust, style=boxed, tabsize=2]
async fn main() {
	let motor_a = ....  // setup boilerplate
	let motor_b = ....
	let motor_c = ....

	let motor_a = motor_a.maintain()  // future
	let motor_b = motor_b.maintain()  // future
	let motor_c = motor_c.maintain()  // future

	join!(motor_a, motor_b, motor_c)
}
\end{lstlisting}
\end{minipage}

Since the join function blocks asynchronously until all its arguments resolve to a value and the maintain futures never resolve the join ends up driving the entire program. To add a tasks simply add it to the join statement. For example to test the motors I include a test future that switches the motors speed every few seconds. It is composed of normal synchrounus code to tell the motors to change their speed target and an asynchronous sleep function.

Non of the functions used in \textit{motor maintain} may block. They can only 'block' asynchronously using \textit{await}. This requires async building blocks, these are provided by the \textit{Embassy}\cite{embassy} project.

\subsection{Safe and Unsafe}

\subsection{Hardware}

\section{Results}
\label{sec:results}

%\input{content/intro}
%\section{Theory}
%\input{content/theory}
%\section{Implementation}
%\input{content/impl}
%\section{Results}
%\input{content/res}
%\section{Conclusion}
%\input{content/conclusion}
	% not everything non blocking need sleeps etc
	% how to do blocking on embedded (can not put on other thread)

\clearpage
\appendix
%\section{Run Instructions}
% \input{content/run_instructions}
\printbibliography

\end{document}

\pagestyle{scrheadings} % Show chapter titles as headings
\cleardoublepage % Avoids problems with pdfbookmark
